\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{fancyhdr}
\usepackage{datetime2}
\usepackage{parskip}
\usepackage{lipsum}

\pagestyle{fancy}
\fancyhf{}

\lhead{eriro331, micso554}
\rhead{\today} % yyyy-mm-dd
\setlength{\headheight}{15pt}

\cfoot{\thepage}

\begin{document}

\begin{center}
    \Huge
    \textbf{TDDC17 - AI}

    \vspace{0.3cm}
    \Large
    Erik Rönmark
    Michael Sörsäter

    \vspace{0.7cm}
    \textbf{Lab3 report}
\end{center}

\section{Part 2}
\subsection{Question 5}
\textbf{What is the risk of melt-down in the power plant during a day if no observations have been made? What if there is icy weather?}

The risk of meltdown is 2.578\% \\
If there is icy weather the probability increases to 3.472\%

\textbf{Suppose that both warning sensors indicate failure. What is the risk of a meltdown in that case? Compare this result with the risk of a melt-down when there is an actual pump failure and water leak. What is the difference? The answers must be expressed as conditional probabilities of the observed variables, P(Meltdown|...).}

P(Meltdown $|$ (PumpFailureWarning, WaterLeakWarning)) = 14.535\% \\
P(Meltdown $|$ (PumpFailure, WaterLeak)) = 20\%

\textbf{The conditional probabilities for the stochastic variables are often estimated by repeated experiments or observations. Why is it sometimes very difficult to get accurate numbers for these? What conditional probabilites in the model of the plant do you think are difficult or impossible to estimate?}

The problem with stochastic variables are that they change over time. E.g. IcyWeather is very hard to model because it can depend on things like season.

\textbf{Assume that the ``IcyWeather'' variable is changed to a more accurate ``Temperature'' variable instead (don't change your model). What are the different alternatives for the domain of this variable? What will happen with the probability distribution of P(WaterLeak $|$ Temperature) in each alternative?}

True and False is a bad representation of temperature. A better example would be to use ranges. E.g. 0-5,5-10,... The probability distribution of P(WaterLeak $|$ Temperature) will vary depending of the observed range. More narrow ranges will lead to a more accurate model.

\subsection{Question 6}
\textbf{What does a probability table in a Bayesian network represent?}

For each possible combination of the parents probabilities, shows the probability for the different domains.

\textbf{What is a joint probability distribution? Using the chain rule on the structure of the Bayesian network to rewrite the joint distribution as a product of P(child $|$ parent) expressions, calculate manually the particular entry in the joint distribution of P(Meltdown=F, PumpFailureWarning=F, PumpFailure=F, WaterLeakWaring=F, WaterLeak=F, IcyWeather=F). Is this a common state for the nuclear plant to be in?}

It is a probability based on multible variables.

P($\lnot$Meltdown, $\lnot$PumpFailureWarning, $\lnot$PumpFailure, $\lnot$WaterLeakWaring, $\lnot$WaterLeak, $\lnot$IcyWeather) = \\ P($\lnot$Meltdown $|$ $\lnot$PumpFailureWarning, $\lnot$PumpFailure, $\lnot$WaterLeakWaring, $\lnot$WaterLeak, $\lnot$IcyWeather) $*$ \\ P($\lnot$PumpFailureWarning $|$ $\lnot$PumpFailure, $\lnot$WaterLeakWaring, $\lnot$WaterLeak, $\lnot$IcyWeather) $*$ \\ P($\lnot$PumpFailure $|$ $\lnot$WaterLeakWaring, $\lnot$WaterLeak, $\lnot$IcyWeather) $*$ \\ P($\lnot$WaterLeakWaring $|$ $\lnot$WaterLeak, $\lnot$IcyWeather) $*$ \\
P($\lnot$WaterLeak $|$ $\lnot$IcyWeather) $*$ \\ P($\lnot$IcyWeather) $=$ \\
$0.999 * 0.95 * 0.9 * 0.95 * 0.9 * 0.95 \approx 69.4\%$

Yes it is a common state to be in.

\textbf{What is the probability of a meltdown if you know that there is both a water leak and a pump failure? Would knowing the state of any other variable matter? Explain your reasoning!}

The probability is 20\%. No other variables matter since the warnings have no link to Meltdown and IcyWeather only affects Waterleak, but Waterleak is already observed as true.

\textbf{Calculate manually the probability of a meltdown when you happen to know that PumpFailureWarning=F, WaterLeak=F, WaterLeakWarning=F and IcyWeather=F but you are not really sure about a pump failure.}

P(Meltdown $|$ $\lnot$PumpFailureWarning, $\lnot$WaterLeakWaring, $\lnot$WaterLeak, \\ $\lnot$IcyWeather) = \\
$\alpha *$ P(Meltdown, $\lnot$PumpFailureWarning, $\lnot$WaterLeakWaring, $\lnot$WaterLeak, $\lnot$IcyWeather) = \\
$\alpha *$ P(Meltdown $|$ $\lnot$PumpFailureWarning, $\lnot$WaterLeakWaring, $\lnot$WaterLeak, $\lnot$IcyWeather) $*$ \\ P($\lnot$PumpFailureWarning $|$ $\lnot$WaterLeakWaring, $\lnot$WaterLeak, $\lnot$IcyWeather) $*$ \\ P($\lnot$WaterLeakWaring $|$ $\lnot$WaterLeak, $\lnot$IcyWeather) $*$ \\ P($\lnot$WaterLeak $|$ $\lnot$IcyWeather) $*$ \\ P($\lnot$IcyWeather) = \\

$\alpha = \frac{1}{(0.865 * 0.95 * 0.9 * 0.95)} \approx 1.42329$

P(Meltdown $|$ ...) = $\alpha * 0.00272 * 0.865 * 0.95 * 0.9 * 0.95 = \alpha * 0.0019110618 \approx 0.00272$

P($\lnot$Meltdown $|$ ...) = $\alpha * 0.99728 * 0.865 * 0.95 * 0.9 * 0.95 = \alpha * 0.7006851882 \approx 0.99728$

Control: $0.00272 + 0.99728 = 1$ %that shit is lit

The probability for a meltdown is 0.272\%

\section {Part 3}
\subsection{Question 2}
\textbf{During the lunch break, the owner tries to show off for his employees by demonstrating the many features of his car stereo. To everyone's disappointment, it doesn't work. How did the owner's chances of surviving the day change after this observation?}

Before P(Survives) = 0.99001 \\
After P(Survives) = 0.98116

\textbf{The owner buys a new bicycle that he brings to work every day. The bicycle has the following properties: \\
\begin{itemize}
\item P(bicycle$\_$works) = 0.9 \\
\item P(survives $|$ $\lnot$moves $\land$ melt-down $\land$ bicycle$\_$works) = 0.6 \\
\item P(survives $|$ moves $\land$ melt-down $\land$ bicycle$\_$works) = 0.9  \\
\end{itemize}
How does the bicycle change the owner's chances of survival?}

Before P(Survives) = 0.99001 \\
After P(Survives) = 0.99505

\textbf{It is possible to model any function in propositional logic with Bayesian Networks. What does this fact say about the complexity of exact inference in Bayesian Networks? What alternatives are there to exact inference?}

% Old answer:  \\
% Yes it is possible. However, the complexity grows exponentially in worst case. The alternative is approximative inference.

New answer in swedish: \\
I ett Bayesiskt nätverk växer komplexiteten väldigt snabbt vilket gör att det blir väldigt prestandakrävande. I värsta fall fås komplexiteten, för n variabler, $O(n*2^n)$. I Propositional logic är komplexiteten $O(2^n)$ vilket gör att komplexiteten i Bayesiska nätverk är värre.

Alternativen till exact inference är att arbeta med approximationer. En approximationsmetod är Monte Carlo simuleringar där slumpen används. Monte Carlo kan ge en väldigt bra approximation!

\section {Part 4}
\subsection{Question 2}
\textbf{The owner had an idea that instead of employing a safety person, to replace the pump with a better one. Is it possible, in your model, to compensate for the lack of Mr H.S.'s expertise with a better pump?}

In our model, if the new pump have the probability of failure set to 2\% or less it is better to replace the pump than employing Mr H.S.

HS:
P(Survives) = 0.99706

Better Pump (2\% failure):
P(Survives) = 0.99725


\textbf{Mr H.S. fell asleep on one of the plant's couches. When he wakes up he hears someone scream: "There is one or more warning signals beeping in your control room!". Mr H.S. realizes that he does not have time to fix the error before it is to late (we can assume that he wasn't in the control room at all). What is the chance of survival for Mr H.S. if he has a car with the same properties as the owner?}

If there is one or more warnings the probability is: \\
P(HS survives) = 0.6128\%

\textbf{What unrealistic assumptions do you make when creating a Bayesian Network model of a person?}

Our model of a person have so few variables, to accurate model a person you need a lot of variables. Humans are random.
% https://www.youtube.com/watch?v=GJY8jJkDoMY

\textbf{Describe how you would model a more dynamic world where for example the "IcyWeather" is more likely to be true the next day if it was true the day before. You only have to consider a limited sequence of days.}

Markov chains using states with transition depending on previous states.

\end{document}
